GxHash: A High-Throughput, Non-Cryptographic Hashing Algorithm Leveraging Modern CPU Capabilities

### 2023-09-19 article/article.tex
Add some references

\documentclass[10pt]{article} 
\usepackage[margin=0.8in]{geometry}
\usepackage{amsmath}
\usepackage{hyperref} 
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multicol}

@@

\colorbox{yellow}{TODO}

\begin{lstlisting}[language=Rust, style=boxed]
use std::hint::black_box;

fn sum1(x: u32) -> u32 {
    let mut sum = x;
    for i in 0..1000 {
        sum = black_box(sum + i);
    }
    sum
}
\end{lstlisting}

@@

Let \( h: \{0,1\}^{n_b \times s_b} \to \{0,1\}^{s_h} \) be a cryptographic hash function where:

\begin{itemize}
    \item Input \( M \) is a binary message composed of \( n_b \) blocks of size \( s_b \),
    such as \( M = M_1 \parallel M_2 \parallel \ldots \parallel M_{n_b} \).
    \item Output is a binary hash of size \( s_h \).
\end{itemize}

@@

In the \textbf{Merkle–Damgård} construction, given:
\begin{itemize}
    \item The compression function \( f: \{0,1\}^{s_b} \times \{0,1\}^{s_b} \to \{0,1\}^{s_b} \)
    \item The pipe function \( g: \{0,1\}^{s_b} \to \{0,1\}^{s_h} \)
    \item A zero-filled vector \( 0^{s_b} \) of length \( s_b \)
\end{itemize}
The hash function \( h \) can be expressed as:
\begin{align*}
    h(M) &= g\left( f(\ldots f(f(0^{s_b}, M_1), M_2) \ldots, M_{n_b}) \right)
\end{align*}

@@

\href{https://www.researchgate.net/publication/322094216_Merkle-Damgard_Construction_Method_and_Alternatives_A_Review}{Alternatives}.

\section{High ILP Construction} \label{highilp}

@@

@@

>>>>

\documentclass[10pt]{article} 
\usepackage[margin=0.8in]{geometry}
\usepackage{amsmath}
\usepackage{hyperref} 
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{graphicx}

@@

As a software engineer at Equativ, a company specializing in high-performance AdServing backends that handle billions of auctions daily,
I face unique challenges in maximizing throughput while minimizing latency. In this high-stakes environment, every millisecond counts,
and the performance of underlying data structures becomes critically important. We heavily rely on in-memory caches and other hash-based data structures,
making the efficiency of hashing algorithms a non-trivial concern in our system's overall performance.\\\\
While diving into the theory of hashing out of both necessity and intellectual curiosity, I discovered that existing hashing algorithms,
including those built on well-known constructions like Merkle–Damgård, are not optimized to exploit the full capabilities of modern general-purpose CPUs.
These CPUs offer advanced features such as Single Instruction, Multiple Data (SIMD) and Instruction-Level Parallelism (ILP), which remain largely untapped
by current hashing methods.\\\\
The challenge of creating a faster, more efficient hashing algorithm became not just a professional necessity but also a personal quest.
It was both challenging and exhilarating to delve into hashing theory and experiment with new approaches. The result is what I believe to be the fastest
non-cryptographic hashing algorithm developed to date.\\\\
The primary motivation behind this research is to bridge the existing performance gap by designing a hashing algorithm that fully leverages SIMD and ILP
capabilities. The aim is to achieve an order-of-magnitude improvement in hashing speed, thereby revolutionizing the efficiency of various applications,
from databases to real-time analytics and beyond.\\\\
In summary, this work is driven by both the practical needs of my professional environment and a personal passion for pushing the boundaries of what is
technically possible in the realm of hashing algorithms.

@@

The field of hashing algorithms has been a subject of extensive research and development, tracing its roots back to foundational architectures like
the \textbf{Merkle–Damgård} Construction\cite{merkle}\cite{damgard}, introduced in 1989. Over the years, this area has seen a plethora of innovations, each attempting to address various
aspects of hashing—be it collision resistance, distribution uniformity, or computational efficiency. While cryptographic hashing has often been the focal
point of research, non-cryptographic hashing algorithms have also garnered attention for their utility in data structures like hashmaps and caches.

In this section, we will explore some of the seminal works and recent advancements in the realm of hashing algorithms to contextualize
our research.

@@

The Merkle–Damgård construction serves as the foundational architecture for many existing hash functions.
It operates by breaking down an input into fixed-size blocks, which are then processed sequentially through a compression
function. The output of each block feeds into the next, culminating in a final, fixed-size hash.\\\\

Let \( h: \{0,1\}^{n_b \times s_b} \to \{0,1\}^{s_h} \) be a cryptographic hash function where: 
\begin{itemize}
    \item Input \( M \) is a binary message composed of \( n_b \) blocks of size \( s_b \),
    such as \( M = M_1 \parallel M_2 \parallel \ldots \parallel M_{n_b} \).
    \item Output is a binary hash of size \( s_h \).
\end{itemize}

The Merkle–Damgård construction can be formalized like so. Given:
\begin{itemize}
    \item The compression function \( f: \{0,1\}^{s_b} \times \{0,1\}^{s_b} \to \{0,1\}^{s_b} \)
    \item The pipe function \( g: \{0,1\}^{s_b} \to \{0,1\}^{s_h} \)
    \item A zero-filled vector \( 0^{s_b} \) of length \( s_b \)
\end{itemize}
The hash function \( h \) can be expressed as:
\begin{align*}
    h(M) &= g\left( f(\ldots f(f(0^{s_b}, M_1), M_2) \ldots, M_{n_b}) \right)
\end{align*}

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{linear-construction.png}
\caption{Merkle–Damgård Construction Overview}
\label{fig:linear-construction}
\end{figure}

@@

\cite{merkle_damgard_alternatives_review}

\clearpage
\section{High ILP Construction} \label{highilp}

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{laned-construction.png}
\caption{Laned Construction Overview}
\label{fig:linear-construction}
\end{figure}

@@

\bibliography{references}
\bibliographystyle{plain}
